Sensitivity Analysis

### 2023-10-11 SFE_article_Sensitivity_Analysis/Article_Model_Parameters.tex
11.10

\begin{keywords}
Supercritical extraction \sep Parameter estimation \sep Mathematical modelling
\end{keywords}

@@

\subsection{Governing equations} \label{CH:Governing_equations_chapter}
The governing equation \footnote{For the sake of clarity of the process model, different colors have been used in the equations to indicate: 
	{\color{red}control variables},
	{\color{blue}state variables},
	{\color{orange}variables} and
	{\color{magenta}parameters}.} for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the Appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}. Quasi-one-dimensional flow is a fluid flow characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as an irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates are given by Equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_3}. The derivation of these Equations are presented in Appendix \ref{CH: Gouverning equations}.

{\footnotesize
	\begin{align}
		\label{EQ: CompressibleEuler_1}
		\cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) v \right)}{\partial z} &= 0 \\
		\cfrac{\partial \left( {\color{orange}\rho_f} v {\color{magenta}A_f}(z) \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) v^2 \right)}{\partial z} &= -{\color{magenta}A_f}(z) \cfrac{\partial {\color{blue}P}}{\partial z} \label{EQ: CompressibleEuler_2} \\
		\cfrac{\partial \left( {\color{orange}\rho_f} {\color{blue}e} {\color{magenta}A_f}(z) \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f} {\color{magenta}A_f}(z) v {\color{blue}e}\right)}{\partial z} &= -{\color{blue}P}\cfrac{\left( {\color{magenta}A_f}(z) v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial {\color{blue}T}}{\partial z} \right)   
		\label{EQ: CompressibleEuler_3}
	\end{align}  
}

where ${\color{orange}\rho_f}$ is the density of the fluid, ${\color{magenta}A_f}(z)$ is the function which describe change of the cross-section, $v$ is the velocity, ${\color{blue}P}$ is the total pressure, ${\color{blue}e}$ is the internal energy of the fluid, $t$ is time and $z$ is the spacial direction.

Based on governing equations, the small discontinuity (defined as $\delta$) in flow properties, shown in Figure \ref{fig: Discontinuity_slow_flow}, can be analysed. The analysis follows the work of \citet{Schreier1982}.

\begin{figure}[!h]
	\centering
	\resizebox{0.95\columnwidth}{!}{%
		\begin{tikzpicture}[]
			\draw (0,2) -- (9,2);	% Top line
			\draw (0,0) -- (9,0); 	% Bottom line
			\draw (4.5,0) -- (4.5,2); 	% Bottom line
			\node at (4,1.5) {${\color{orange}\rho_f}$};
			\node at (5.5,1.5) {${\color{orange}\rho_f}+\delta{\color{orange}\rho_f}$};
			\node at (4,1.0) {${\color{blue}P}$};
			\node at (5.5,1.0) {${\color{blue}P}+\delta {\color{blue}P}$};
			\node at (4,0.5) {${\color{blue}T}$};
			\node at (5.5,0.5) {${\color{blue}T}+\delta {\color{blue}T}$};
			\node at (1,1.0) {$v \rightarrow$};
			\node at (8,1.0) {$v + \delta v \rightarrow$};
	\end{tikzpicture} }
	\caption{Small discontinuity in one-dimensional flow}
	\label{fig: Discontinuity_slow_flow}
\end{figure}

The discontinuity is presumed to be at rest relative, and the balance equations become

{\footnotesize
	\begin{align*}
		&{\color{orange}\rho_f} \delta v + v \delta {\color{orange}\rho_f} + \delta {\color{orange}\rho_f} \delta v = 0 \\
		&\delta {\color{blue}P} = \delta v \delta {\color{orange}\rho_f}
	\end{align*}
}

These relations are equally valid if the two regions are separated by a region of finite width rather than a discontinuity.

{\footnotesize
	\begin{equation*}
		\lim_{{\color{orange}\rho_f} v \rightarrow 0} {\color{orange}\rho_f} \delta v + v \delta {\color{orange}\rho_f} + \delta {\color{orange}\rho_f} \delta v = 0 / \delta {\color{orange}\rho_f} \rightarrow \cfrac{d v}{d {\color{orange}\rho_f}} = - \cfrac{v}{{\color{orange}\rho_f}}
	\end{equation*}
}

By combining the momentum equation with the above equation, we get

{\footnotesize
	\begin{equation} \label{EQ: Pressure_Velocity}
		\cfrac{d v}{d {\color{orange}\rho_f}} = - \cfrac{d v}{d{\color{blue}P}} \cfrac{d {\color{blue}P}}{d {\color{orange}\rho_f}} = -\cfrac{1}{\rho v} \cfrac{d{\color{blue}P}}{d{\color{orange}\rho_f}} = -\cfrac{v}{{\color{orange}\rho_f}}
	\end{equation}
}

Suppose the flow is presumed to be isentropic, $d{\color{blue}P}/d{\color{orange}\rho_f} = c^2$, so $v^2=c^2$, where $c$ is the speed of sound. This can be interpreted as a small pressure wave propagating with the speed of sound relative to the flow. Moreover, if the flow velocity is relatively low, all pressure changes are hydrodynamic (due to velocity motion) rather than thermodynamic which leads to $\partial {\color{orange}\rho_f} / \partial {\color{blue}P} \approx 0$. In other words, the small changes in pressure due to flow velocity changes do not change the density. This has a secondary effect -- the speed of sound in the fluid is $\partial {\color{blue}P}/\partial {\color{orange}\rho_f} = \infty$ in this instance. So there is an infinite speed of sound, which makes the equations elliptic in nature. It can be deduced that at the isothermal conditions, the density in the system propagates with the same speed as pressure since they are both connected through the equation of state. The details of the Mach-number analysis are presented in Appendix \ref{CH:Low_Mach_chapter}.

@@

\clearpage \appendix \label{appendix}
\section{Appendix} 
\subsection{Thermodynamic}
\subfile{Sections/Thermodynamic_details}

\subsection{Governing equations}
\subfile{Sections/Gouverning_equation_derivation}

\subsection{Low Mach number expansion} \label{CH:Low_Mach_chapter}
%\subfile{Sections/Low_mach_number_expansion}
\subfile{Sections/Low_mach_number_expansion_imp}

\subsection{Bayes theorem} \label{CH: Bayes}
\subfile{Sections/Bayes_Theorem}

\subsection{Cardano's Formula} \label{CH: Cardano}
\subfile{Sections/Cardano}

\subsection{Initial and boundary conditions} \label{CH: IC_BC}
\subfile{Sections/IC_BC}

@@

\subsection{Solid density measurement} \label{CH: Solid_Density_Measurment}

\begin{figure}[!h]
	\centering 
	\includegraphics[trim=2cm 6cm 4cm 0cm, clip,width=\columnwidth]{Sections/ultraReportT5.pdf}
	\caption{The result of solid density measurement}
\end{figure}

{\footnotesize
	\begin{equation*}
		\rho_s^{ave} = \frac{1.2585+1.2582+1.2561+1.2546+1.2555}{5} = 1.25658 [g/cc]
	\end{equation*}
}

\subsection{Porosity Calculations} \label{CH: Porosity}
\subfile{Sections/Porosity}

>>>>

\begin{keywords}
Supercritical extraction \sep Sensitivity analysis \sep Mathematical modelling
\end{keywords}

@@

\subsection{Governing equations} \label{CH:Governing_equations_chapter}
\subfile{Sections/Gouverning_equation}

@@

%\clearpage \appendix \label{appendix}
%\section{Appendix} 
%\subsection{Thermodynamic}
%\subfile{Sections/Qubic_EOS}

%\subsection{Governing equations}
%\subfile{Sections/Gouverning_equation_derivation}

%\subsection{Low Mach number expansion} \label{CH:Low_Mach_chapter}
%\subfile{Sections/Low_mach_number_expansion}
%\subfile{Sections/Low_mach_number_expansion_imp}

%\subsection{Bayes theorem} \label{CH: Bayes}
%\subfile{Sections/Bayes_Theorem}

%\subsection{Cardano's Formula} \label{CH: Cardano}
%\subfile{Sections/Cardano}

%\subsection{Initial and boundary conditions} \label{CH: IC_BC}
%\subfile{Sections/IC_BC}

@@

%\subsection{Solid density measurement} \label{CH: Solid_Density_Measurment}

%\begin{figure}[!h]
%	\centering 
%	\includegraphics[trim=2cm 6cm 4cm 0cm, clip,width=\columnwidth]{Sections/ultraReportT5.pdf}
%	\caption{The result of solid density measurement}
%\end{figure}

%{\footnotesize
%	\begin{equation*}
%		\rho_s^{ave} = \frac{1.2585+1.2582+1.2561+1.2546+1.2555}{5} = 1.25658 [g/cc]
%	\end{equation*} }

%\subsection{Porosity Calculations} \label{CH: Porosity}
%\subfile{Sections/Porosity}

### 2023-10-11 SFE_article_Sensitivity_Analysis/Sections/Gouverning_equation.tex
11.10

% ---------------------------------------------------------------
% Preamble
% ---------------------------------------------------------------
%\documentclass[a4paper,fleqn,longmktitle]{cas-sc}
\documentclass[a4paper,fleqn]{cas-dc}
%\documentclass[a4paper]{cas-dc}
%\documentclass[a4paper]{cas-sc}
% ---------------------------------------------------------------
% Make margins bigger to fit annotations. Use 1, 2 and 3. TO be removed later
%\paperwidth=\dimexpr \paperwidth + 6cm\relax
%\oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
%\evensidemargin=\dimexpr\evensidemargin + 3cm\relax
%\marginparwidth=\dimexpr \marginparwidth + 3cm\relax
% -------------------------------------------------------------------- 
% Packages
% --------------------------------------------------------------------
% Figure packages
\usepackage{graphicx}
\usepackage{float}
\restylefloat{table}
\usepackage{adjustbox}
% Text, input, formatting, and language-related packages
\usepackage[T1]{fontenc}
\usepackage{subcaption}

\usepackage{csvsimple}

% TODO package
\usepackage[bordercolor=gray!20,backgroundcolor=blue!10,linecolor=black,textsize=footnotesize,textwidth=1in]{todonotes}
\setlength{\marginparwidth}{1in}
% \usepackage[utf8]{inputenc}
% \usepackage[nomath]{lmodern}

% Margin and formatting specifications
%\usepackage[authoryear]{natbib}
\usepackage[sort]{natbib}
\setcitestyle{square,numbers}

 %\bibliographystyle{cas-model2-names}

\usepackage{setspace}
\usepackage{subfiles} % Best loaded last in the preamble

% \usepackage[authoryear,longnamesfirst]{natbib}

% Math packages
\usepackage{amsmath, amsthm, amssymb, amsfonts, bm, nccmath, mathdots, mathtools, bigints, ulem}

\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{shapes.geometric,angles,quotes,calc}

\usepackage{placeins}

\usepackage[final]{pdfpages}

% --------------------------------------------------------------------
% Packages Configurations
\usepackage{enumitem}
% --------------------------------------------------------------------
% (General) General configurations and fixes
\AtBeginDocument{\setlength{\FullWidth}{\textwidth}}	% Solves els-cas caption positioning issue
\setlength{\parindent}{20pt}
%\doublespacing
% --------------------------------------------------------------------
% Other Definitions
% --------------------------------------------------------------------
\graphicspath{{Figures/}}
% --------------------------------------------------------------------
% Environments
% --------------------------------------------------------------------
% ...

% --------------------------------------------------------------------
% Commands
% --------------------------------------------------------------------

% ==============================================================
% ========================== DOCUMENT ==========================
% ==============================================================
\begin{document} 
%  --------------------------------------------------------------------

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Parameter estimation}                      
\shorttitle{Parameter estimation}

\shortauthors{OS, PO}

\author[1]{Oliwer Sliczniuk}[orcid=0000-0003-2593-5956]
\ead{oliwer.sliczniuk@aalto.fi}
\cormark[1]
\credit{a}

\author[1]{Pekka Oinas}[orcid=0000-0002-0183-5558]
\credit{b}

%\author[1]{Francesco Corona}[orcid=0000-0002-3615-1359]
%\credit{c}

\address[1]{Aalto University, School of Chemical Engineering, Espoo, 02150, Finland}
%\address[2]{2}

\cortext[cor1]{Corresponding author}

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
This study aimed to investigate the supercritical extraction process of the caraway oil from caraway seeds. A distributed-parameter model is used to describe the fluid-solid extraction process with carbon dioxide as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent are estimated from the Peng-Robinson equation of state. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient. The missing model parameters and the initial state are obtained by solving the maximum likelihood estimation based on experimental yield data under the normal error assumption. A set of laboratory experiments was performed in a partially filled extractor with a fixed bed operated under multiple constant operating conditions: $40~^\circ C$ / $200$ bar, $50~^\circ C$ / $200$ bar, $40~^\circ C$ / $300$ bar, and $50~^\circ C$ / $300$ bar.

\end{abstract}

\begin{keywords}
Supercritical extraction \sep Parameter estimation \sep Mathematical modelling
\end{keywords}

% ===================================================
% TITLE
% ===================================================
\maketitle

% ===================================================
% Section: Introduction
% ===================================================\section{Introduction}

\section{Introduction}
%\subfile{Sections/introduction}
\subfile{Sections/introduction_imp}

\section{Materials and methods} \label{CH: Materials and methods}

\subsection{Supercritical fluids} \label{CH: Thermodynamic}
%\subfile{Sections/Thermo}
\subfile{Sections/Thermo_imp}

%\clearpage
%\subfile{Sections/Materials_and_methods}

\subsection{Governing equations} \label{CH:Governing_equations_chapter}
%\footnote{For the sake of clarity of the process model, different colors have been used in the equations to indicate: 
%	{\color{red}control variables},
%	{\color{black}state variables},
%	{\color{black}variables} and
%	{\color{black}parameters}.} 
	The governing equation for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the Appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}. Quasi-one-dimensional flow is a fluid flow characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as an irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

@@

\subsection{Extraction model} \label{CH: Extraction_model}
\subfile{Sections/Model}

%\newpage
%\section{Bayes theorem} \label{CH: Bayes}
%\subfile{Sections/Bayes_Theorem}
%\clearpage

\subsection{Parameter estimation} \label{CH: Parameter_estimation}
\subfile{Sections/Parameter_estimation}

\subsection{Experimental work}
\subfile{Sections/Experiments}

\section{Results}
\subfile{Sections/Results}

\section{Conclusions} \label{CH: Conclusion}

This study presented basic aspects regarding the laboratory work for obtaining caraway extracts using supercritical carbon dioxide at different operating conditions. A first-principle process model was developed based on governing equations. As a result of the literature review, the extraction kinetic model was selected and combined with the general mass balance equations. The values of unknown parameters in the process model were obtained by parameter estimation. The yield data from the experiments were compared against the yield generated by the model. The IPOPT solver was used to solve the maximum likelihood estimation. The process model was found to have good agreement with experimental yield curves. Later, the parameters found for each experiment were combined to obtain more general relationships. The correlations for diffusion and decay coefficients are presented as a function of fluid density or Reynolds number. It should be noted that obtained correlations have been prepared with a limited amount of data. These correlations can be used to study qualitatively the impact of different operating conditions on the extraction yield by different methods, such as sensitivity analysis. Moreover, the generalized model could be useful in finding the optimal operating conditions with respect to the process economy.

% ===================================================
% Bibliography
% ===================================================
%% Loading bibliography style file
\clearpage
%\bibliographystyle{model1-num-names}
\bibliographystyle{unsrtnat}
\bibliography{mybibfile}

\clearpage \appendix \label{appendix}
\section{Appendix} 
\subsection{Thermodynamic}
\subfile{Sections/Qubic_EOS}
%\subfile{Sections/Thermodynamic_details}

\subsection{Governing equations}
\subfile{Sections/Gouverning_equation_derivation}

%\subsection{Low Mach number expansion} \label{CH:Low_Mach_chapter}
%\subfile{Sections/Low_mach_number_expansion}
%\subfile{Sections/Low_mach_number_expansion_imp}

\subsection{Cardano's Formula} \label{CH: Cardano}
\subfile{Sections/Cardano}

\subsection{Initial and boundary conditions} \label{CH: IC_BC}
\subfile{Sections/IC_BC}

%\subsection{Bayes theorem} \label{CH: Bayes}
%\subfile{Sections/Bayes_Theorem}

\subsection{Maximum likelihood} \label{CH: ML}
\subfile{Sections/Likelihood}

\subsection{Solid density measurement} \label{CH: Solid_Density_Measurment}

Figure \ref{fig:density_cal} shows results of density measurement performed with pycnometer.

\begin{figure}[!h]
	\centering 
	\includegraphics[trim=2cm 6cm 4cm 0cm, clip,width=\columnwidth]{Sections/ultraReportT5.pdf}
	\caption{The result of solid density measurement}
	\label{fig:density_cal}
\end{figure}

{\footnotesize
	\begin{equation*}
		\rho_s^{ave} = \frac{1.2585+1.2582+1.2561+1.2546+1.2555}{5} = 1.25658 [g/cc]
	\end{equation*}
}

\subsection{Porosity calculations} \label{CH: Porosity}
\subfile{Sections/Porosity}

%\subsection{Concentration profiles} \label{CH: Profiles}
%\subfile{Sections/Profiles}

%\subsection{Solution of the parameter estimation starting from another initial guess} \label{CH: Profiles_2}
%\subfile{Sections/Parameter_estimation_solution_from_another_IG}

>>>>

\documentclass[../Article_Model_Parameters.tex]{subfiles}
\graphicspath{{\subfix{../Figures/}}}
\begin{document}

	%\footnote{For the sake of clarity of the process model, different colors have been used in the equations to indicate: 
		%	{\color{red}control variables},
		%	{\color{black}state variables},
		%	{\color{black}variables} and
		%	{\color{black}parameters}.} 
	The governing equation for a quasi-one-dimensional compressible flow in Cartesian coordinates can be found in the Appendix \ref{CH: Gouverning equations} and in the work of \citet{Anderson1995}. Quasi-one-dimensional flow is a fluid flow characterized by the assumption that the flow properties remain uniform across any given cross-section of the flow. This assumption is made when there is a variation in the cross-sectional area of the flow channel, such as an irregular shape or partial filling of an extractor. In such cases, the flow is considered to be quasi-one-dimensional because the velocity and other flow properties are assumed to vary only in the direction of flow.

@@

### 2023-10-11 SFE_article_Sensitivity_Analysis/Sections/Likelihood.tex
11.10

	To find the maximum likelihood estimates, we maximize the joint probability density function, or likelihood function, denoted as $p(\theta | y(t_1),y(t_2),...,y(t_n))$, where $\theta$ represents the parameters and $y(t_1),y(t_2),...,y(t_n)$ represent the observed data. The conditions at the maximum can be refined by incorporating prior information. The posterior probability density function $p(\theta|y)$ can be expressed as the ratio of two probability densities (\ref{EQ: Probabiltiy_independet_b}) using the continuous variable analogue of Bayes' theorem (details in Appendix \ref{CH: Bayes}). In such a case the posteriori distribution is given by Equation \ref{EQ: Posterior_distribution}.

@@

	The numerator of the right-hand side od Equation \ref{EQ: Posterior_distribution} using Equation \ref{EQ: Probabiltiy_dependet_a} becomes

>>>>

	To find the maximum likelihood estimates, we maximize the joint probability density function, or likelihood function, denoted as $p(\theta | y(t_1),y(t_2),...,y(t_n))$, where $\theta$ represents the parameters and $y(t_1),y(t_2),...,y(t_n)$ represent the observed data. The conditions at the maximum can be refined by incorporating prior information. The posterior probability density function $p(\theta|y)$ can be expressed as the ratio of two probability densities %(\ref{EQ: Probabiltiy_independet_b})
	 using the continuous variable analogue of Bayes' theorem %(details in Appendix \ref{CH: Bayes})
	 . In such a case the posteriori distribution is given by Equation \ref{EQ: Posterior_distribution}.

@@

	The numerator of the right-hand side od Equation \ref{EQ: Posterior_distribution} becomes %using Equation \ref{EQ: Probabiltiy_dependet_a} becomes
